### PML - Writeup
#### Summary of steps
The basic steps in involved are:
   * Loading the training set - pml-training.csv
   * Cleaning up the dataset to remove the columns/predictors with near zero variance and where there are lot of missing values or NAs. near zero variance predictors take unique values across samples, are non-informative as well as break the model.
   * From the training set above, create a model training set and a cross-validation set in the ratio of 70:30 percent.
   * Using the model training set above, train a model using RandomForest
   * Run varImp() on the generated model to checkout the significance of predictors
   * Run the cross-validation set against the generated model to predict the model accuracy.
   * Check the predictions from this cross-validated set using a confusion matrix.
   * If the accuracy is > 0.9; run the real test set from pml-testing.csv

#### Expected out of sample error
The out of sample or out of bag error for the model is 0.21% given by model$finalModel:
OOB estimate of  error rate: 0.21%.
The confusion matrix for cross-validation shows a percentage error of 14/(1674+1133+1022+963+1079)*100 = 0.23%. 

With a > 0.9 accuracy, 95% of the out of sample data should lie within the CI generated by the model. It is appropriately shown below for the cross-validation confusion matrix.

#### Confusion Matrix and Statistics on cross-validation data

<!-- html table generated in R 3.1.1 by xtable 1.7-4 package -->
<!-- Mon Sep 22 00:48:54 2014 -->
<table border=1>
<tr> <th>  </th> <th> A </th> <th> B </th> <th> C </th> <th> D </th> <th> E </th>  </tr>
  <tr> <td align="right"> A </td> <td align="right"> 1674 </td> <td align="right">   6 </td> <td align="right">   0 </td> <td align="right">   0 </td> <td align="right">   0 </td> </tr>
  <tr> <td align="right"> B </td> <td align="right">   0 </td> <td align="right"> 1133 </td> <td align="right">   4 </td> <td align="right">   0 </td> <td align="right">   0 </td> </tr>
  <tr> <td align="right"> C </td> <td align="right">   0 </td> <td align="right">   0 </td> <td align="right"> 1022 </td> <td align="right">   0 </td> <td align="right">   0 </td> </tr>
  <tr> <td align="right"> D </td> <td align="right">   0 </td> <td align="right">   0 </td> <td align="right">   0 </td> <td align="right"> 963 </td> <td align="right">   3 </td> </tr>
  <tr> <td align="right"> E </td> <td align="right">   0 </td> <td align="right">   0 </td> <td align="right">   0 </td> <td align="right">   1 </td> <td align="right"> 1079 </td> </tr>
   </table>

Overall Statistics
                                         
               Accuracy : 0.9976         
                 95% CI : (0.996, 0.9987)
    No Information Rate : 0.2845         
    P-Value [Acc > NIR] : < 2.2e-16
    
#### Implementation
```{r model,echo=TRUE,cache=TRUE}
#df <- read.csv("/Users/mridul/coursera/PREDMACHLEARN/pml-training.csv")
#explore <- describe(df)
#library(caret)
#removecols <- nearZeroVar(df)
#nearzerodf <- df[,-removecols]
#cleandf <- subset(nearzerodf, select=-c(X,user_name,raw_timestamp_part_1,raw_timestamp_part_2,cvtd_timestamp,var_accel_forearm,amplitude_pitch_forearm,min_pitch_forearm,max_picth_forearm,var_yaw_dumbbell,stddev_yaw_dumbbell,avg_yaw_dumbbell,var_pitch_dumbbell,stddev_pitch_dumbbell,avg_pitch_dumbbell,var_roll_dumbbell,stddev_roll_dumbbell,avg_roll_dumbbell,var_accel_dumbbell,amplitude_pitch_dumbbell,amplitude_roll_dumbbell,min_pitch_dumbbell,min_roll_dumbbell,max_picth_dumbbell,max_roll_dumbbell,max_roll_belt,max_picth_belt,min_roll_belt,min_pitch_belt,amplitude_roll_belt,amplitude_pitch_belt,var_total_accel_belt, avg_roll_belt,stddev_roll_belt, var_roll_belt,avg_pitch_belt, stddev_pitch_belt,var_pitch_belt,avg_yaw_belt,stddev_yaw_belt,var_yaw_belt,var_accel_arm,max_picth_arm,max_yaw_arm,min_yaw_arm,amplitude_yaw_arm))

#traindf <- cleandf
#select <- createDataPartition(traindf$classe, p = 0.7, list = FALSE)
#modelTrain <- traindf[select,]
#modelTest <- traindf[-select,]
#model <- train(classe ~ ., data=modelTrain, method="rf", verbose=FALSE)
```
```{r predictions,echo=TRUE}
#predictions <- predict(model,modelTest)
#confusionMatrix(predictions,modelTest$classe)

#dftest <- read.csv("/Users/mridul/coursera/PREDMACHLEARN/pml-testing.csv")
#cleanset <- dftest[,names(cleandf)[1:53]]
#clean <- cbind(cleanset,dftest$problem_id)
#Do not show the answer for pml-testing.csv. So comment out the following; though it works with 100% accuracy on pml-testing.csv
##realpredictions <- predict(model,clean)
##summary(realpredictions)
```
